Search for Spatially Extended \fermi-LAT Sources Using Two Years of Flight Data

### 2011-08-28 extended_catalog/paper/extended_catalog.tex
collaboration meeting presentation

\author{
J.~Lande\altaffilmark{2},
S.~Funk\altaffilmark{2},
M.~Ackermann (???),
A.~Allafort\altaffilmark{2},
M.-H.~Grondin\altaffilmark{53},
% Keith Bechtol
% Marshall Roth?
% Marianne Lemoine-Goumard
% Francesco Giordano
% Toby Burnett
% Jean Ballet
% Matthew Kerr
\ldots
\altaffiltext{2}{W. W. Hansen Experimental Physics Laboratory, Kavli Institute for Particle Astrophysics and Cosmology, Department of Physics and SLAC National Accelerator Laboratory, Stanford University, Stanford, CA 94305, USA}
\altaffiltext{53}{Institut f\"ur Astronomie und Astrophysik, Universit\"at T\"ubingen, D 72076 T\"ubingen, Germany}
}

@@

A number of astrophysical source classes including supernova remnants
(SNRs), pulsar wind nebulae (PWNe), molecular clouds, normal galaxies,
and galaxy clusters are expected to be spatially resolvable at
\gev energies.  Dark matter satellites are also hypothesized to
be spatially extended.  The Large Area Telescope
(LAT) on the {\em \fermi Gamma-ray Space Telescope} (\fermi) has
detected five SNRs which are significantly extended
at \gev energies: IC443, W28, W44, W51C, and RX\,J1713.7$-$3946
\citep{ic443,w28,w44,w51c,rx_j1713_lat}. In addition, three extended
PWNe were detected as being extended: MSH\,15$-$52, Vela~X, and
HESS\,J1825$-$137 \citep{msh1552,velax,fermi_hess_j1825}. Two
close-by galaxies, the Large and Small Magalenic Clouds, and one
radio galaxy, Centarus A, were spatially resolved at \gev energies
\citep{lmc,smc,cen_a_lat}.  A number of additional sources detected
at \gev energies are positionally coincident with sources that exhibit
extension at other wavelengths. These sources could be expected to be spatially
resolvable by the LAT at \gev energies.

@@

The LAT is a pair conversion telescope that been surveying the
$\gamma$-ray sky since 2008 June.  The LAT has broad energy coverage
(20 \mev to $>300$ \gev), wide field of view ($\sim 2.4 \sr$), and large
effective area ($\sim 8000\ \cm^2$ at $>1 \gev$).  
Using one year of all-sky surveying data, the LAT Collaboration published
a catalog of 1451 sources significantly detected at \gev energies called 1FGL \citep{first_cat}.
Using two years of data, a second catalog called 2FGL reported on 1873
sources \citep{second_cat}.
The counterparts of many of these sources can be spatially resolved
when observed at other frequencies but detecting the spatial extension
of these sources at \gev energies is difficult because the size of the
point-spread function (PSF) of the LAT is comparable to the typical size
of many of these sources.

@@

\hl{
Figure~\ref{all_sensitivity} shows the flux threshold
as a function of source extension for different background levels 
($1\times$,
$10\times$, 
and $100\times$ the nominal background), different spectral indices, and
two different energy bands (1 \gev to 100 \gev and 10 \gev to 100 \gev).  The detection
threshold is higher for sources in regions of higher background.  For a
fixed 1 \gev to 100 \gev or 10 \gev to 100 \gev flux, the LAT's detection threshold
only weakly depends upon the spectral index.  This effect is most
pronounced when using only photons with energies above 10 \gev.  Overlaid
on Figure~\ref{all_sensitivity} are the LAT detected extended sources
that will be discussed in Sections~\ref{validate_known}
and \ref{new_ext_srcs_section}.  The extension thresholds are tabulated
in Table~\ref{all_sensitivity_table}.

@@

We first present on our analysis of the twelve extended sources
included in 2FGL.
\citep{second_cat}.
Six extended SNRs were included in 2FGL: IC443, W28, W30, W44, W51C, and the
Cygnus Loop \citep{ic443,w28,w44,w51c,cygnus_loop_lat}.\footnote{A
detailed publication by the LAT
Collaboration about W30 is still in preparation.}
\gev emission from W30 was also studied in \cite{castro_and_slane_2010}.
Using 1 \gev to 100 \gev photons, our analysis significantly detected the
extension of all six SNRs.
% 2 galaxies

@@

The source 2FGL\,J0823.0$-$4246 was found 
to have an 
extension of $0\fdg37\pm0\fdg03_\stat\pm0\fdg02_\sys$ 
with an extension
significance of $\tsext=46.3$
using 1 \gev to 100 \gev photons.  This source was fit to a position of
$(l,b)=(260\fdg32,-3\fdg28)$.  
This source is coincident with the one-year
catalog source 1FGL\,J0823.3$-$4248.
\hl{
Figure~\ref{1FGL_J0823.3-4248} shows a
counts map of this source
and Figure~\ref{snr_seds} shows the LAT spectrum of it.
}
To get a good fit of this source, we removed from our background
model the nearby catalog
sources 2FGL\,J0823.4$-$4305 and 2FGL\,J0821.0$-$4254 which are part of the
extended source.  These modifications are
shown in Figure~\ref{1FGL_J0823.3-4248}.

\hl{
This extended source is spatially coincident with the middle-aged
SNR Puppis A.  Puppis A has been studied in detail in radio
(\cite{puppis_a_vla}, and references therein) and  X-ray 
(\cite{rosat_puppis_a,suzaku_puppis_a}, and references therein)
Mosaic \rosat observations of Puppis A produced the highest
resolution X-ray image of this source
and contours corresponding to this
image are overlaid on Figure~\ref{1FGL_J0823.3-4248} and match the
inferred size at \gev energies \citep{rosat_puppis_a}.
Puppis A's distance was estimated at 2.2 kpc \citep{reynoso_1995,reynoso_2003}
which leads to a 1 \gev to 100 \gev luminosity of $\sim 3\times 10^{34}$ ergs/\s.

No molecular clouds have been observed directly adjacent to Puppis A 
\citep{co_eastern_puppis_a}.
This is similar to the LAT detected SNR the Cyngus Loop \citep{cygnus_loop_lat}.
The luminosity of Puppis A is also smaller than that of 
other SNRs believed to interact with molecular clouds
\citep{w51c,ic443,w44,w28,w49b_lat}.
The \gev emission could therefore be explained by the interaction
of accelerated particles by interstellar matter or fields, similar
to the emission mechanism suggested for the Cygnus Loop SNR
\citep{cygnus_loop_lat}.
}

@@

\hl{
This source is in a region of remarkably complicated diffuse emission.
Even though it is $16\deg$ from the Galactic plane, this source is on
top of the core of the Ophiuchus molecular cloud which contains massive
star-forming regions that are bright in the infrared.  The region also has
abundant molecular and atomic gas traced by CO and HI but also plenty of
dark gas found only by its association with dust emission
\citep{isabelle_dark_gass}. Embedded star-forming regions make it even
more challenging to measure the column density of dust.  Infared and 
${}^{12}\text{CO}$ ($J=1\rightarrow 0$)
contours are overlaid on Figure~\ref{1FGL_J1628.6-2419c}. There is good
spatial correlation with the \gev emission \citep{iras_rho_ophiuci,co_rho_ophiuci}
so this source might represent an inadequacy in the diffuse
$\gamma$-ray model.
}

@@

2FGL\,J1615.2$-$5138 is spatially coincident with the extended
\tev source HESS\,J1616$-$508 \citep{hess_plane_survey}.  In
Figure~\ref{1FGL_J1613.6-5100c}, contours of HESS\,J1616$-$508 are overlaid
on 2FGL\,J1615.0$-$5051.  The H.E.S.S. experiment measured an
extension $0.136\pm 0.008$ when fitting this source with an elliptical
2D Gaussian surface brightness.  This size corresponds to a 68\% containment
radius of $\rsixeight=0\fdg21\pm0\fdg01$. This size is comparable to the LAT
size $\rsixeight=0\fdg27\pm0\fdg03$ (see Section~\ref{compare_source_size}).
Figure~\ref{hess_seds} shows that the LAT spectrum of 2FGL\,J1615.0$-$5051
connects smoothly to the H.E.S.S spectrum of HESS\,J1616$-$508.

@@

This source is spatially coincident with the extended
\tev source HESS\,J1614$-$518 \citep{hess_plane_survey}. In
Figure~\ref{1FGL_J1613.6-5100c}, contours of HESS\,J1614$-$518 are overlaid
on 2FGL\,J1615.2$-$5138.  The H.E.S.S. experiment measured a 2D Gaussian
extension of $\sigma=0\fdg23\pm0\fdg02$ and $\sigma=0.15\pm0.02$
in the semi-major and semi-minor axis. This size corresponds
to a 68\% containment size of $\rsixeight=0\fdg35\pm0\fdg03$
and $0.23\pm0.03$.  This elliptical size matches the LAT size
$\rsixeight=0\fdg35\pm0\fdg03$.  Figure~\ref{hess_seds} shows
that the LAT spectrum of 2FGL\,J1615.2$-$5138 connects smoothly to
the H.E.S.S spectrum of HESS\,J1614$-$518.  Further data collected by
H.E.S.S. in 2007 helped to resolve a double peaked structure in the
H.E.S.S. data but no spectral variation across this source, suggesting
that the emission is not the confusion of physically separate sources
\citep{closer_look_hess_j1614-518}.  The source was also detected by
CANGAROO-III \citep{cangaroo_j1614-518}.

@@

This extended source is spatially coincident with the extended
\tev source HESS\,J1632$-$478 \citep{hess_plane_survey}.
In Figure~\ref{1FGL_J1632.9-4802c}, contours of  HESS\,J1632$-$478
are overlaid on 2FGL\,J1635.4$-$4717c.  H.E.S.S measured a 
extension of $\sigma=0.21\pm0.05$ and $0.06\pm0.04$ along the
semi-major and semi-minor axes when fitting this source with an
elliptical 2D Gaussian surface brightness.  This corresponds to a 68\%
containment size $\rsixeight=0\fdg31\pm0\fdg08$ and $0\fdg09\pm0\fdg06$
along the semi-major and semi-minor axis. This size is consistent with
the LAT size $\rsixeight=0\fdg36\pm0\fdg04$.  Figure~\ref{hess_seds}
shows that the LAT spectrum of 2FGL\,J1635.4$-$4717c connects smoothly to
the H.E.S.S spectrum of HESS\,J1632$-$478.

\hl{
\cite{hess_plane_survey} argued that HESS\,J1632$-$478
is positionally coincident with the hard X-ray source
IGR\,J1632-4751 observed by INTEGRAL, \xmmnewton, and \asca
\citep{Igr_J16320-4751_circ,xmm_newton_IGR_J16320-4751,asca_plane_survey},
but this source is suspected to be an galactic X-Ray Binary so the $\gamma$-ray
extension disfavors the association.  Further observations by
\xmmnewton reveal extended emission of size $\sim32\arcsec\times15\arcsec$
and inside point-like emission coincident with the peak
H.E.S.S. emission \citep{hess_j1632_478_xmm_newton}.  They found
in archival MGPS-2 data a spatially coincident extended radio source
\citep{most_survey_galactic_plane}.  The positional match argues for a
single synchrotron processes producing X-ray and \gev and \tev radiation,
likely due to a PWNe.  The increased size at \tev energies
compared
to X-ray energies has previously been observed in several aging PWNe
including HESS\,J1825$-$137 \citep{hess_j1825_xmm_newton,hess_j1825_hess},
HESS\,J1640$-$465 \citep{hess_plane_survey,xmm_newton_hess_j_1640-466},
and Vela X \citep{vela_x_rosat,vela_x_hess}.  This can be explained by
a different synchrotron cooling time for the electrons producing X-rays
and \tev $\gamma$-rays.
}

@@

This source is spatially coincident with the \tev source HESS\,J1837$-$069
\citep{hess_plane_survey}.  In Figure~\ref{1FGL_J1837.5-0659c}, contours
of HESS\,J1837$-$069 are overlaid on 2FGL\,J1837.3$-$0700c. H.E.S.S. measured
an extension of $\sigma=0.12\pm0.02$ and $0.05\pm0.02$
along the semi-major and semi-minor axis when fitting this source
with an elliptical 2D Gaussian surface brightness.  This corresponds
to a 68\% containment radius of $\rsixeight=0\fdg18\pm0\fdg03$ and
$0\fdg08\pm0\fdg03$ along the semi-major and semi-minor axis. The
size is comparable to LAT which fit a 68\% containment radius of
$\rsixeight=0\fdg29\pm0\fdg07$.  the size difference is not significant
(less than 2 sigma).  Figure~\ref{hess_seds} shows that the LAT spectrum
connects smoothly to the H.E.S.S spectrum of HESS\,J1837$-$069.

\hl{
HESS\,J1837$-$069 is coincident
with the hard and steady X-ray source AX\,J1838.0$-$0655
\citep{einstein_galactic_plane_survey,hard_x-ray_asca,integral_AX_J1838.0-0655,swift_follow_up,pulsations_HESS_J1837-069,suzaku_HESS_J1837-069}.
This source was discovery by RXTE to be a pulsar sufficiently
luminous to power the \tev emission.  AX\,J1838.0$-$0655 was spatially
resolved by \chandra to be a bright point source surrounded by
a $\sim2\arcmin$ nebula \citep{pulsations_HESS_J1837-069} and so the
$\gamma$-ray emission may be powered by this pulsar.  A second
X-ray point source AX\,J1837.3$-$0652 is in the region of HESS\,J1837$-$069
\citep{hard_x-ray_asca,swift_follow_up,pulsations_HESS_J1837-069,suzaku_HESS_J1837-069}.
It was also resolved into point-like and diffuse component
although no pulsations have yet been detected from it
\citep{pulsations_HESS_J1837-069}.  If AX\,J1838.0$-$0655 is a pulsar/PWN
powering HESS\,J1837$-$069, some of the \tev emission may also come from
AX\,J1837.3$-$0652.
}

@@

The source 2FGL\,J2021.5+4026 was found 
to have an extension of $0\fdg59\pm0\fdg03_\stat\pm0\fdg02_\sys$
with an extension significance of $\tsext=116.4$
using 10 \gev to 100 \gev photons.  This source was
fit to a position of $(l,b)=(78\fdg18,2\fdg19)$.  This source
is coincident with the one-year catalog source 1FGL\,J2020.0+4049.
\hl{
Figure~\ref{1FGL_J2020.0+4049} shows a counts map of this source.
and Figure~\ref{snr_seds} shows that the LAT spectrum of it.
}

@@

\hl{
This extended sources is coincident with 2CG\,078+01 and
3EG\,J2020+4017 detected by COS B and EGRET respectively
\citep{second_cos_b_catalog,third_egret_catalog}.  This source is
spatially coincident with the $\gamma$-Cygni SNR and has been speculated
to be be the interaction of accelerated particles in the SNR interacting
with dense molecular clouds \citep{pollock_1985,gaisser_1998}.  But this
association was disfavored when the LAT \gev emission from this source
was detected to be pulsed 
(PSR\,J2021+4026,\cite{first_lat_pulsar_cat}).

@@

\hl{
Twelve extended sources were included in 2FGL.  Using two years of
LAT data and a new analysis method, we presented the detection of
nine additional extended sources. We also reanalyze the spatial
shape of the twelve extended sources in 2FGL.  The 21 extended
LAT sources are located close to the Galactic plane and plotted in
Figure~\ref{allsky_extended_sources}.  Most of the extended sources are
indeed expected to be of Galactic origin as the distance of extragalactic
sources (with the exception of the local group Galaxies) is typically too
large to resolve them in $\gamma$-rays.

@@

Figure~\ref{gev_vs_tev_histogram} compares the size of the 21 extended
LAT sources to the 42 extended H.E.S.S. sources.\footnote{The 
\tev extension of
the 42 extended H.E.S.S. sources comes from the H.E.S.S. Source
Catalog. The H.E.S.S. source catalog can be found at \url{http://www.mpi-hd.mpg.de/hfm/HESS/pages/home/sources/}.}
Because of the large
field of view and all-sky coverage, the LAT can more easily measure
larger sources including the LMC and SMC.  On the other hand, the smaller
angular resolution of Air Cherenkov detectors allows them to measure a
population of extended sources below our sensitivity limit (currently at
about $\sim0\fdg2$).  \fermi has a five year nominal mission lifetime with
a goal of ten years of operation.  As Figure~\ref{time_sensitivity} shows,
the LAT's low background at high energies will allow it's sensitivity to
these smaller sources to improve by a factor greater than the square root
of the exposure.  With increased exposure, the LAT will likely begin to
detect and resolve many smaller \tev sources.

Figure~\ref{compare_index_2FGL} compares the spectral index
of LAT detected extended sources and of all sources in 2FGL. This,
and Table~\ref{known_extended_sources} and~\ref{new_ext_srcs_table},
show that the LAT observes a population of hard extended sources at
energies above 10 \gev.  Figure~\ref{hess_seds} shows that the
\gev spectral of
four of these sources (2FGL\,J1615.0$-$5051, 2FGL\,J1615.2$-$5138,
2FGL\,J1632.4$-$4753c, and 2FGL\,J1837.3$-$0700c)  
connect smoothly to their \tev counterparts. This is also true of Vela
Jr., HESS\,J1825$-$137 \citep{fermi_hess_j1825}, RX\,J1713.7$-$3946
\citep{rx_j1713_lat}.  It is likely that the \gev and \tev emission
from these sources originates from the same population of high-energy
particles.

Many of the \tev detected extended sources now seen at \gev energies are
currently unidentified and further multi-wavelength follow-up observations
will be necessary to understand these particle accelerators. But \gev
observations of these \tev sources in many cases can help to constrain
handronic and leptonic models of these sources and, especially if their
spectra can be extended to lower energies, help to understand the emission
mechanisms producing these high energy $\gamma$-rays.  
}

>>>>

\author{
J.~Lande\altaffilmark{2},
S.~Funk\altaffilmark{2},
M.~Ackermann (???),
%A.~Allafort\altaffilmark{2},
%M.-H.~Grondin\altaffilmark{53},
% Keith Bechtol
% Marshall Roth?
% Marianne Lemoine-Goumard
% Francesco Giordano
% Toby Burnett
% Jean Ballet
% Matthew Kerr
\ldots
\altaffiltext{2}{W. W. Hansen Experimental Physics Laboratory, Kavli Institute for Particle Astrophysics and Cosmology, Department of Physics and SLAC National Accelerator Laboratory, Stanford University, Stanford, CA 94305, USA}
\altaffiltext{53}{Institut f\"ur Astronomie und Astrophysik, Universit\"at T\"ubingen, D 72076 T\"ubingen, Germany}
}

@@

A number of astrophysical source classes including supernova remnants
(SNRs), pulsar wind nebulae (PWNe), molecular clouds, normal galaxies,
and galaxy clusters are expected to be spatially resolvable at
\gev energies.  Dark matter satellites are also hypothesized to
be spatially extended.  The Large Area Telescope
(LAT) on the {\em \fermi Gamma-ray Space Telescope} (\fermi) has
detected seven SNRs which are significantly extended
at \gev energies: 
W51C, W30, 
IC443, W28, W44, RX\,J1713.7$-$3946,
and the Cygnus Loop
\citep{w51c,castro_and_slane_2010,ic443,w28,w44,rx_j1713_lat,cygnus_loop_lat}. In addition, three extended
PWNe were detected as being extended: MSH\,15$-$52, Vela~X, and
HESS\,J1825$-$137 \citep{msh1552,velax,fermi_hess_j1825}. Two
close-by galaxies, the Large and Small Magalenic Clouds, and one
radio galaxy, Centarus A, were spatially resolved at \gev energies
\citep{lmc,smc,cen_a_lat}.  A number of additional sources detected
at \gev energies are positionally coincident with sources that exhibit
extension at other wavelengths, large enough to be spatially
resolvable by the LAT at \gev energies.

@@

The LAT is a pair conversion telescope that has been surveying the
$\gamma$-ray sky since 2008 June.  The LAT has broad energy coverage
(20 \mev to $>300$ \gev), wide field of view ($\sim 2.4 \sr$), and large
effective area ($\sim 8000\ \cm^2$ at $>1 \gev$).  
Using one year of all-sky surveying data, the LAT Collaboration published
a catalog of 1451 sources significantly detected at \gev energies called 1FGL \citep{first_cat}.
Using two years of data, a second catalog called 2FGL reported on 1873
sources \citep{second_cat}.
The counterparts of many of these sources can be spatially resolved
when observed at other frequencies but detecting the spatial extension
of these sources at \gev energies is difficult because the size of the
point-spread function (PSF) of the LAT is comparable to the typical size
of many of these sources.

@@

\hl{
Figure~\ref{all_sensitivity} shows the flux threshold as a function of
source extension for different background levels ($1\times$, $10\times$,
and $100\times$ the nominal background), different spectral indices,
and two different energy bands (1 \gev to 100 \gev and 10 \gev to
100 \gev).  The detection threshold is higher for sources in regions of
higher background.  When studying sources only at energies above 1 \gev,
the LAT's detection threshold (defined as the 1 \gev to 100 \gev flux at
which $\langle\tsext\rangle=16$) depends less strongly on the source's
spectral index. The index dependence of the detection threshold is
even weaker when considering only photons with energies above 10 \gev.
Overlaid on Figure~\ref{all_sensitivity} are the LAT detected extended
sources that will be discussed in Sections~\ref{validate_known} and
\ref{new_ext_srcs_section}.  The extension thresholds are tabulated in
Table~\ref{all_sensitivity_table}.

@@

We first present on our analysis of the twelve extended sources
included in 2FGL.  \citep{second_cat}.  Six extended SNRs were
included in 2FGL: W51C, IC443, W28, W30, W44, and the Cygnus Loop
\citep{w51c,ic443,w28,w44,cygnus_loop_lat}.\footnote{A detailed
publication by the LAT Collaboration about W30 is still in preparation.}
\gev emission from W30 was also studied in \cite{castro_and_slane_2010}.
Using 1 \gev to 100 \gev photons, our analysis significantly detected
the extension of all six SNRs.
% 2 galaxies

@@

The source 2FGL\,J0823.0$-$4246 was found 
to have an 
extension of $0\fdg37\pm0\fdg03_\stat\pm0\fdg02_\sys$ 
with an extension
significance of $\tsext=46.3$
using 1 \gev to 100 \gev photons.  This source was fit to a position of
$(l,b)=(260\fdg32,-3\fdg28)$.  
This source is coincident with the one-year
catalog source 1FGL\,J0823.3$-$4248.
\hl{
Figure~\ref{1FGL_J0823.3-4248} shows a
counts map of this source
and Figure~\ref{snr_seds} shows its spectrum.
}
To get a good fit of this source, we removed from our background
model the nearby catalog
sources 2FGL\,J0823.4$-$4305 and 2FGL\,J0821.0$-$4254 which are part of the
extended source.  These modifications are
shown in Figure~\ref{1FGL_J0823.3-4248}.

\hl{
This extended source is spatially coincident with the
middle-aged SNR Puppis A.  Puppis A has been studied in detail
in radio (\cite{puppis_a_vla}, and references therein) and  X-ray
(\cite{rosat_puppis_a,suzaku_puppis_a}, and references therein) Mosaic
\rosat observations of Puppis A produced the highest resolution X-ray
image of this source and contours corresponding to this image are
overlaid on Figure~\ref{1FGL_J0823.3-4248} and match the inferred
size at \gev energies \citep{rosat_puppis_a}.  Puppis A's distance
was estimated at 2.2 kpc \citep{reynoso_1995,reynoso_2003} which leads
to a 1 \gev to 100 \gev luminosity of $\sim 3\times 10^{34}$ ergs/\s.
No molecular clouds have been observed directly adjacent to Puppis A
\citep{co_eastern_puppis_a}.  This is similar to the LAT detected Cyngus
Loop SNR \citep{cygnus_loop_lat}.  The luminosity of Puppis A is also
smaller than that of other SNRs believed to interact with molecular clouds
\citep{w51c,ic443,w44,w28,w49b_lat}.
\citep{cygnus_loop_lat}.
}

@@

\hl{
This source is in a region of remarkably complicated diffuse emission.
Even though it is $16\deg$ from the Galactic plane, this source is on
top of the core of the Ophiuchus molecular cloud which contains massive
star-forming regions that are bright in the infrared.  The region also has
abundant molecular and atomic gas traced by CO and HI but also plenty of
dark gas found only by its association with dust emission
\citep{isabelle_dark_gass}. Embedded star-forming regions make it even
more challenging to measure the column density of dust.  Infared and 
${}^{12}\text{CO}$ ($J=1\rightarrow 0$)
contours are overlaid on Figure~\ref{1FGL_J1628.6-2419c}. There is good
spatial correlation with the \gev emission \citep{iras_rho_ophiuci,co_rho_ophiuci}
so this source might 
represent $\gamma$-ray emission from the interactions of cosmic rays with
interstellar gas which has not been accounted for in the LAT diffuse
emission model.
}

@@

2FGL\,J1615.2$-$5138 is spatially coincident with the extended
\tev source HESS\,J1616$-$508 \citep{hess_plane_survey}.  In
Figure~\ref{1FGL_J1613.6-5100c}, contours of HESS\,J1616$-$508 are overlaid
on 2FGL\,J1615.0$-$5051.  The H.E.S.S. experiment measured an
extension $0.136\pm 0.008$ when fitting this source with an elliptical
2D Gaussian surface brightness.  This size corresponds to a 68\% containment
radius of $\rsixeight=0\fdg21\pm0\fdg01$. This size is comparable to the LAT
size $\rsixeight=0\fdg27\pm0\fdg03$ (see Section~\ref{compare_source_size}).
Figure~\ref{hess_seds} shows that the spectrum of 2FGL\,J1615.0$-$5051
at \gev energies
connects smoothly to the spectrum of HESS\,J1616$-$508 at \tev
energies.

@@

This source is spatially coincident with the extended
\tev source HESS\,J1614$-$518 \citep{hess_plane_survey}. In
Figure~\ref{1FGL_J1613.6-5100c}, contours of HESS\,J1614$-$518 are overlaid
on 2FGL\,J1615.2$-$5138.  The H.E.S.S. experiment measured a 2D Gaussian
extension of $\sigma=0\fdg23\pm0\fdg02$ and $\sigma=0.15\pm0.02$
in the semi-major and semi-minor axis. This size corresponds
to a 68\% containment size of $\rsixeight=0\fdg35\pm0\fdg03$
and $0.23\pm0.03$.  This elliptical size matches the LAT size
$\rsixeight=0\fdg35\pm0\fdg03$.  Figure~\ref{hess_seds} shows
that the spectrum of 2FGL\,J1615.2$-$5138 at \gev energies
connects smoothly to
the spectrum of HESS\,J1614$-$518 at \tev energies.  Further data collected by
H.E.S.S. in 2007 helped to resolve a double peaked structure in the
H.E.S.S. data but no spectral variation across this source, suggesting
that the emission is not the confusion of physically separate sources
\citep{closer_look_hess_j1614-518}.  The source was also detected by
CANGAROO-III \citep{cangaroo_j1614-518}.

@@

This extended source is spatially coincident with the extended
\tev source HESS\,J1632$-$478 \citep{hess_plane_survey}.
In Figure~\ref{1FGL_J1632.9-4802c}, contours of  HESS\,J1632$-$478
are overlaid on 2FGL\,J1635.4$-$4717c.  H.E.S.S measured a 
extension of $\sigma=0.21\pm0.05$ and $0.06\pm0.04$ along the
semi-major and semi-minor axes when fitting this source with an
elliptical 2D Gaussian surface brightness.  This corresponds to a 68\%
containment size $\rsixeight=0\fdg31\pm0\fdg08$ and $0\fdg09\pm0\fdg06$
along the semi-major and semi-minor axis. This size is consistent with
the LAT size $\rsixeight=0\fdg36\pm0\fdg04$.  Figure~\ref{hess_seds}
shows that the spectrum of 2FGL\,J1635.4$-$4717c at \gev energies
connects smoothly to
the spectrum of HESS\,J1632$-$478 at \tev energies.

\hl{
\cite{hess_plane_survey} argued that HESS\,J1632$-$478
is positionally coincident with the hard X-ray source
IGR\,J1632-4751 observed by INTEGRAL, \xmmnewton, and \asca
\citep{Igr_J16320-4751_circ,xmm_newton_IGR_J16320-4751,asca_plane_survey},
but this source is suspected to be a galactic X-Ray Binary so the $\gamma$-ray
extension disfavors the association.  Further observations by
\xmmnewton reveal extended emission of size $\sim32\arcsec\times15\arcsec$
and inside point-like emission coincident with the peak
H.E.S.S. emission \citep{hess_j1632_478_xmm_newton}.  They found
in archival MGPS-2 data a spatially coincident extended radio source
\citep{most_survey_galactic_plane}.  The positional match argues for a
single synchrotron process producing X-ray and \gev and \tev radiation,
likely due to a PWNe.  The increased size at \tev energies
compared
to X-ray energies has previously been observed in several aging PWNe
including HESS\,J1825$-$137 \citep{hess_j1825_xmm_newton,hess_j1825_hess},
HESS\,J1640$-$465 \citep{hess_plane_survey,xmm_newton_hess_j_1640-466},
and Vela X \citep{vela_x_rosat,vela_x_hess}.  This can be explained by
a different synchrotron cooling time for the electrons producing X-rays
and \tev $\gamma$-rays.
}

@@

This source is spatially coincident with the \tev source HESS\,J1837$-$069
\citep{hess_plane_survey}.  In Figure~\ref{1FGL_J1837.5-0659c}, contours
of HESS\,J1837$-$069 are overlaid on 2FGL\,J1837.3$-$0700c. H.E.S.S. measured
an extension of $\sigma=0.12\pm0.02$ and $0.05\pm0.02$
along the semi-major and semi-minor axis when fitting this source
with an elliptical 2D Gaussian surface brightness.  This corresponds
to a 68\% containment radius of $\rsixeight=0\fdg18\pm0\fdg03$ and
$0\fdg08\pm0\fdg03$ along the semi-major and semi-minor axis. The
size is comparable to LAT which fit a 68\% containment radius of
$\rsixeight=0\fdg29\pm0\fdg07$.  the size difference is not significant
(less than 2 sigma).  Figure~\ref{hess_seds} shows that the spectrum
of 2FGL\,J1837.3$-$0700c at \gev energies
connects smoothly to the spectrum of HESS\,J1837$-$069
at \tev energies.

\hl{
HESS\,J1837$-$069 is coincident
with the hard and steady X-ray source AX\,J1838.0$-$0655
\citep{einstein_galactic_plane_survey,hard_x-ray_asca,integral_AX_J1838.0-0655,swift_follow_up,pulsations_HESS_J1837-069,suzaku_HESS_J1837-069}.
This source was discovered by RXTE to be a pulsar sufficiently
luminous to power the \tev emission.  AX\,J1838.0$-$0655 was spatially
resolved by \chandra to be a bright point source surrounded by
a $\sim2\arcmin$ nebula \citep{pulsations_HESS_J1837-069} and so the
$\gamma$-ray emission may be powered by this pulsar.  A second
X-ray point source AX\,J1837.3$-$0652 is in the region of HESS\,J1837$-$069
\citep{hard_x-ray_asca,swift_follow_up,pulsations_HESS_J1837-069,suzaku_HESS_J1837-069}.
It was also resolved into point-like and diffuse component
although no pulsations have yet been detected from it
\citep{pulsations_HESS_J1837-069}.  If AX\,J1838.0$-$0655 is a pulsar/PWN
powering HESS\,J1837$-$069, some of the \tev emission may also come from
AX\,J1837.3$-$0652.
}

@@

The source 2FGL\,J2021.5+4026 was found to have an extension of
$0\fdg59\pm0\fdg03_\stat\pm0\fdg02_\sys$ with an extension significance
of $\tsext=116.4$ using 10 \gev to 100 \gev photons.  This source
was fit to a position of $(l,b)=(78\fdg18,2\fdg19)$.  This source is
coincident with the one-year catalog source 1FGL\,J2020.0+4049. \hl{
Figure~\ref{1FGL_J2020.0+4049} shows a counts map of this source
and Figure~\ref{snr_seds} shows its spectrum.  
}

@@

\hl{
This extended sources is coincident with 2CG\,078+01 and
3EG\,J2020+4017 detected by COS B and EGRET respectively
\citep{second_cos_b_catalog,third_egret_catalog}.  This source is
spatially coincident with the $\gamma$-Cygni SNR and has been speculated
to originate from 
the interaction of accelerated particles in the SNR interacting
with dense molecular clouds \citep{pollock_1985,gaisser_1998}.  But this
association was disfavored when the LAT \gev emission from this source
was detected to be pulsed 
(PSR\,J2021+4026,\cite{first_lat_pulsar_cat}).

@@

\hl{
Twelve extended sources were included in 2FGL.  Using two years of
LAT data and a new analysis method, we presented the detection of
nine additional extended sources. We also reanalyze the spatial
shape of the twelve extended sources in 2FGL.  The 21 extended
LAT sources are located close to the Galactic plane 
and their localizations shown in
Figure~\ref{allsky_extended_sources}.  Most of the extended sources are
indeed expected to be of Galactic origin as the distance of extragalactic
sources (with the exception of the local group Galaxies) is typically too
large to resolve them in $\gamma$-rays.

@@

Figure~\ref{gev_vs_tev_histogram} compares the size of the 21 extended
LAT sources to the 42 extended H.E.S.S. sources.\footnote{The 
\tev extension of
the 42 extended H.E.S.S. sources comes from the H.E.S.S. Source
Catalog. The H.E.S.S. source catalog can be found at \url{http://www.mpi-hd.mpg.de/hfm/HESS/pages/home/sources/}.}
Because of the large
field of view and all-sky coverage, the LAT can more easily measure
larger sources including the LMC and SMC.  On the other hand, the 
better
angular resolution of Air Cherenkov detectors allows them to measure a
population of extended sources below our sensitivity limit (currently at
about $\sim0\fdg2$).  \fermi has a five year nominal mission lifetime with
a goal of ten years of operation.  As Figure~\ref{time_sensitivity} shows,
the LAT's low background at high energies will allow it's sensitivity to
these smaller sources to improve by a factor greater than the square root
of the exposure.  With increased exposure, the LAT will likely begin to
detect and resolve many smaller \tev sources.

Figure~\ref{compare_index_2FGL} compares the spectral index
of LAT detected extended sources and of all sources in 2FGL. This,
and Table~\ref{known_extended_sources} and~\ref{new_ext_srcs_table},
show that the LAT observes a population of hard extended sources at
energies above 10 \gev.  Figure~\ref{hess_seds} shows that the
spectra of four of these sources (2FGL\,J1615.0$-$5051, 2FGL\,J1615.2$-$5138,
2FGL\,J1632.4$-$4753c, and 2FGL\,J1837.3$-$0700c)  
at \gev energies
connect smoothly to their H.E.S.S. counterparts at \tev
energies. This is also true of Vela
Jr., HESS\,J1825$-$137 \citep{fermi_hess_j1825}, RX\,J1713.7$-$3946
\citep{rx_j1713_lat}.  It is likely that the \gev and \tev emission
from these sources originates from the same population of high-energy
particles.

Many of the \tev detected extended sources now seen at \gev energies are
currently unidentified and further multi-wavelength follow-up observations
will be necessary to understand these particle accelerators.  But the
extension of the spectrum of these \tev sources towards lower energies
by LAT observations can help to determine the origin and nature of the
high-energy emission seen from these sources.
}
